---
title: "Untitled"
author: "Meghan Forrest"
date: "2023-04-11"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(stringr)
library(ggplot2)
library(readxl)
library(table1)
library(qwraps2)
library(ggthemes)
library(ggpubr)
library(gridExtra)
library(datasets)
```


Import Abstraction_master.xlsx. 
```{r}
Abstraction_master <- read.csv(file = url("https://raw.githubusercontent.com/meghanrforrest/SPBG-meta-research-study/main/Abstraction_master.csv"))

# Remove unneeded columns (ID, Startzeit, empty variable, Fertigstellungzeit, E-mail, Name, title, exclusion_details, study_design_other, problems_1, problems_2, problems_3, problems_4, problems_5, abstraction_notes )
data <- 
  Abstraction_master %>% 
  select(!c(1:7, 11, 13, 28, 36, 44, 52, 60, 62)
         )
```


Parse out excluded studies
```{r}
data_exc <- 
  data %>% 
  filter(eligiblity == "No") %>% 
  select(1:3)
```


Parse out included studies
```{r}
data_inc <- 
  data %>% 
  filter(eligiblity == "Yes") %>% 
  select(!3) %>% 
  separate(functional_outcome, into = c("functional_outcome_1", "functional_outcome_2", "functional_outcome_3"),
             sep = ";",   
             fill = "right",   
             remove = FALSE) %>% 
  
# standardize formatting of functional outcomes
  mutate(
    functional_outcome_1 = recode(functional_outcome_1, 
                         '(KOSCHI) King’s Outcome Scale for Childhood Head Injury score' = 'KOSCHI', 
                         'Bayley Scales of Infant Development III' = 'Bayley III', 
                         'Pediatric Stroke Outcome Measure (PSOM) ' = 'PSOM', 
                         'Scales  for  Outcomes  in  PD  (SCOPA)' = "SCOPA", 
                         "Unified  Huntington  Disease  Rating  Scale" = "UHDRS", 
                         "UPDRS" = "UPDRS (composite)", 
                         "Total motor scale (TMS)" = "TMS",
                         "medical research council ( MRC) sum score" = "MRC sum score",
                         "GBS disability score" = "GBS disability scale", 
                         "GBS  disability  scale" = "GBS disability scale",
                         "Pediatric Stroke Outcome Measure (PSOM)" = "PSOM", 
                         "GOS" = "GOS (unextended)")) %>% 
    mutate(
    functional_outcome_2 = recode(functional_outcome_2, 
                         '(KOSCHI) King’s Outcome Scale for Childhood Head Injury score' = 'KOSCHI', 
                         'Bayley Scales of Infant Development III' = 'Bayley III', 
                         'Pediatric Stroke Outcome Measure (PSOM) ' = 'PSOM', 
                         'Scales  for  Outcomes  in  PD  (SCOPA)' = "SCOPA", 
                         "Unified  Huntington  Disease  Rating  Scale" = "UHDRS", 
                         "UPDRS" = "UPDRS (composite)", 
                         "Total motor scale (TMS)" = "TMS",
                         " medical research council ( MRC) sum score" = "MRC sum score",
                         "GBS disability score" = "GBS disability scale", 
                         "GBS  disability  scale" = "GBS disability scale",
                         "Pediatric Stroke Outcome Measure (PSOM)" = "PSOM", 
                         "Multiple Sclerosis Severity Score (MSSS)" = "MSSS", 
                         " total functional capacity (TFC)" = "TFC", 
                         "medical research council ( MRC) sum score" = "MRC sum score", 
                         " Gross Motor Function Classification System (GMFCS)" = "GMFCS", 
                         " GDS" = "GBS disability scale", 
                         "MDS-UPDRS" = "UPDRS (composite)", 
                         " PSOM" = "PSOM", 
                         " UPDRS-III" = "UPDRS-III", 
                         " NIHSS" = "NIHSS", 
                         " Katz index" = "Katz index",
                         " Rasch- Built  Overall Built Disability Score" = "R-ODS", 
                         " PCPC" = "PCPC",
                         "GOS" = "GOS (unextended)"
                         )) %>% 
  mutate(
    functional_outcome_3 = recode(functional_outcome_3,
                                  "Stroke Impact Scale 16" = "SIS-16")) %>% 
    separate(neurological_condition, into = c("neurological_condition_1", "neurological_condition_2"),
             sep = ";",   
             fill = "right",   
             remove = TRUE) %>% 
  
#standardize formatting of neurological conditions
  mutate(neurological_condition_1 = recode(neurological_condition_1,
      "Stroke (all)" = "Stroke (all)", 
      "encephalitis" = "Encephalitis", 
      "GBS" = "Guillain-Barré", 
      "Huntington disease" = "Huntington’s disease", 
      "Huntington Disease" = "Huntington’s disease"
  )) %>% 
  
# turn roc_present into a binary variable 
  mutate(
    roc_present = case_when(
      roc_present == "Neither" ~ 0, 
      roc_present == "Neither;" ~ 0,
      roc_present == "AUC" ~ 1, 
      roc_present == "AUC;" ~ 1, 
      roc_present == "AUC;ROC;" ~ 1, 
      roc_present == "ROC;AUC;" ~ 1
    )) %>% 

# add a variable: whether the study includes an adjusted SPBG
  mutate(
    contains_adj_spbg = case_when(
      adjusted_1 == "Yes" | adjusted_2 == "Yes" | adjusted_3 == "Yes" | adjusted_4 == "Yes" | adjusted_5 == "Yes"~ 1
    ))

data_inc$neurological_condition_1<-gsub(";","",as.character(data_inc$neurological_condition_1))
data_inc$neurological_condition_2<-gsub(";","",as.character(data_inc$neurological_condition_2))

# separate out figure_other_outcome variable (because some studies have two outcomes which use other figures)
data_inc <-
  data_inc %>% 
  separate(figure_other_outcome, into = c("figure_other_outcome_1", "figure_other_outcome_2"),
             sep = ",",   
             fill = "right",   
             remove = TRUE) 


# create variable for how many SPBGs each included record contains
data_inc <- 
  data_inc %>% 
  mutate(
    spbg_count = case_when(
    spbg_present == "No" ~ 0, 
    spbg_present == "Yes" & additional_spbg_1 == "No" ~ 1, 
    spbg_present == "Yes" & additional_spbg_1 == "Yes" & additional_spbg_2 == "No" ~ 2, 
    spbg_present == "Yes" & additional_spbg_1 == "Yes" & additional_spbg_2 == "Yes"  & additional_spbg_3 == "No" ~ 3, 
    spbg_present == "Yes" & additional_spbg_1 == "Yes" & additional_spbg_2 == "Yes" & additional_spbg_3 == "Yes" & additional_spbg_4 == "No" ~ 4, 
    spbg_present == "Yes" & additional_spbg_1 == "Yes" & additional_spbg_2 == "Yes" & additional_spbg_3 == "Yes" & additional_spbg_4 == "Yes" ~ 5
  ))

# create variable for records which contain an adjusted bar graph
data_inc <-
  data_inc %>% 
  mutate(
    contains_adj_spbg = case_when(
      spbg_present == "Yes" & c(adjusted_1 == "Yes" | adjusted_2 == "Yes" | adjusted_3 == "Yes" | adjusted_4 == "Yes" | adjusted_5 == "Yes") ~ 1, 
      spbg_present == "No" ~ 99, 
      spbg_count == 1 & adjusted_1 == "No" ~ 0,
      spbg_count == 2 & c(adjusted_1 == "No" & adjusted_2 == "No") ~ 0, 
      spbg_count == 3 & c(adjusted_1 == "No" & adjusted_2 == "No" & adjusted_3 == "No") ~ 0,
      spbg_count == 4 & c(adjusted_1 == "No" & adjusted_2 == "No" & adjusted_3 == "No" & adjusted_4 == "No") ~ 0,
      spbg_count == 5 & c(adjusted_1 == "No" & adjusted_2 == "No" & adjusted_3 == "No" & adjusted_4 == "No" & adjusted_5 == "No") ~ 0
    ))
  

# MAYBE DELETE THIS create a variable for outcome scales used in SPBG MAYBE DELETE THIS
data_inc <-
  data_inc %>% 
    mutate(
    spbg_outcome  = recode(
      figure_outcome_1,
      "GDS" = "GBS disability scale")) %>% 
  mutate(
    SPBG_outcome = case_when(
      c(figure_outcome_1  == "mRS" | figure_outcome_2 == "mRS" | figure_outcome_3 == "mRS" | data_inc$figure_outcome_4 == "mRS" | data_inc$figure_outcome_5 == "mRS" ) & c(functional_outcome_1 == "mRS" |  functional_outcome_2== "mRS" | functional_outcome_3 == "mRS") ~ "mRS", 
      c(figure_outcome_1  == "GOSE" |data_inc$figure_outcome_2 == "GOSE" | data_inc$figure_outcome_3 == "GOSE" | data_inc$figure_outcome_4 == "GOSE" | data_inc$figure_outcome_5 == "GOSE" ) & c(functional_outcome_1 == "GOSE" |  functional_outcome_2== "GOSE" | functional_outcome_3 == "GOSE") ~ "GOSE", 
      c(figure_outcome_1  == "NIHSS" |data_inc$figure_outcome_2 == "NIHSS" | data_inc$figure_outcome_3 == "NIHSS" | data_inc$figure_outcome_4 == "NIHSS" | data_inc$figure_outcome_5 == "NIHSS" ) & c(functional_outcome_1 == "NIHSS" |  functional_outcome_2== "NIHSS" | functional_outcome_3 == "NIHSS") ~ "NIHSS",
    c(figure_outcome_1  == "PSOM" |data_inc$figure_outcome_2 == "PSOM" | data_inc$figure_outcome_3 == "PSOM" | data_inc$figure_outcome_4 == "PSOM" | data_inc$figure_outcome_5 == "PSOM" ) & c(functional_outcome_1 == "PSOM" |  functional_outcome_2== "PSOM" | functional_outcome_3 == "PSOM") ~ "PSOM", 
    figure_outcome_1  == "GBS disability scale"  & c(functional_outcome_1 == "GBS disability scale" |  functional_outcome_2== "GBS disability scale" | functional_outcome_3 == "GBS disability scale") ~ "GBS disability scale",
    figure_outcome_1  == "RS"  & c(functional_outcome_1 == "RS" |  functional_outcome_2== "RS" | functional_outcome_3 == "RS") ~ "RS"
    ))

# create a variable for whether the study is in primary or secondary analysis
data_inc <-
  data_inc %>% 
  mutate(
    analysis_sample = case_when(
      covariate_intent =="Yes" & roc_present == 0 & effect_estimate != "Beta coefficient(s) only" ~ "primary", 
      covariate_intent =="No" | roc_present == 1 | effect_estimate == "Beta coefficient(s) only" ~ "sensitivity")
    )

```


Create primary and sensitivity analysis dataset
```{r}
data_primary <-
  data_inc %>% 
  filter(analysis_sample == "primary")
  
data_primary_spbg <- 
  data_primary %>% 
  filter(spbg_present == "Yes")

data_sensitivity <- 
  data_inc %>% 
  filter(covariate_intent == "No" | roc_present == 1 | effect_estimate == "Beta coefficient(s) only")

data_sensitivity_spbg <- 
  data_sensitivity %>% 
  filter(spbg_present == "Yes")

```



Create dataset for SPBGs
```{r}
contains_spbg <- 
  data_inc %>% 
  filter(spbg_present == "Yes") %>% 
  select(c(doi, functional_outcome_1, functional_outcome_2, functional_outcome_3, neurological_condition_1, neurological_condition_2, effect_estimate, roc_present, covariate_intent, figure_other_outcome_1, figure_other_outcome_2, spbg_count, contains_adj_spbg, figure_outcome_1, figure_outcome_2, figure_outcome_3, figure_outcome_4, figure_outcome_5, SPBG_outcome)) 



```


Prepare a dataset to make barplots
```{r}
barplot_data <-
  data_inc %>% 
  mutate(
    any_figure_contains = case_when(
     contains_adj_spbg == 0 ~ "Unadjusted SPBG", 
     contains_adj_spbg == 1 ~ "Adjusted SPBG", 
     figure_other == "Yes" & spbg_present == "No" ~ "Other figure (No SPBG)", 
     figure_other == "No" & spbg_present == "No" ~ "No figure")) %>% 
  mutate(
    neurological_condition_1 = recode(
      neurological_condition_1,
      "Ischemic stroke" = "Stroke", 
      "Hemorrhagic stroke" = "Stroke", 
      "Stroke (all)" = "Stroke"
    )) %>% 
  pivot_longer(
    names_to = "which_neuropathology", 
    values_to = "neuropathology", 
    cols = neurological_condition_1:neurological_condition_2) %>% 
  filter(neuropathology != "") 

barplot_data$any_figure_contains <- 
  factor(barplot_data$any_figure_contains, levels = c("No figure", "Other figure (No SPBG)", "Unadjusted SPBG", "Adjusted SPBG"), ordered = TRUE)
```
mutate(
    any_figure_contains = recode(
      any_figure_contains,
      "SPBG" = "Stacked proportional bar graph"))

Make a barplot for stroke vs. other condtions
barplot_data_stroke <- barplot_data
```{r}

barplot_data_stroke <- barplot_data
barplot_data_stroke$neuropathology[barplot_data_stroke$neuropathology %in% names(which(table(barplot_data_stroke$neuropathology) < 100))] = "Other *"

positions_stroke <- c("Other *", "Stroke")


barplot_allneuro <- ggplot(barplot_data_stroke, aes(x= neuropathology, fill = any_figure_contains)) + 
  geom_bar(width = 0.5) + 
  scale_x_discrete(limits = positions_stroke) +
  scale_fill_manual(values = c("#4773a3",  "#a4c3d7", "#FFE082", "grey93"), 
                    limits = c("Adjusted SPBG", "Unadjusted SPBG", "Other figure (No SPBG)", "No figure")) +
  theme(
    panel.background = element_blank(), 
    panel.grid.major.x = element_line(color = "lightgrey",
                                          size = 0.3,
                                          linetype = 5), 
    panel.grid.minor.x = element_line(color = "lightgrey",
                                          size = 0.5,
                                          linetype = 3),
    axis.text.x = element_text(size=  9, face = "bold", colour = "black"), 
    axis.text.y = element_text(size= 9, colour = "black",  margin = margin(r = 5)),
    legend.title=element_blank(),
    legend.key.size = unit(1, 'cm'),
        legend.key.height = unit(.3, 'cm'),
        legend.key.width = unit(.35, 'cm'),
        legend.text = element_text(size = 9), 
    legend.position = c(0.5, 1.0),
    legend.direction="horizontal",
    axis.ticks = element_blank(), 
    axis.title.y=element_blank(), 
    axis.title.x=element_blank(), 
    panel.border = element_rect(colour = "lightgrey", fill=NA, size=0.5), 
    plot.margin = unit(c(0.8,3,0,0), "cm")) +
  coord_flip() +
  geom_text(
    stat= 'count' , aes(label=..count..), 
            position = position_stack(vjust = .5), 
            size = 3, 
            color = 'black') +
  annotate("text", x = 2, y = 203, label = "n = 192", size = 3
           ) +
  annotate("text", x = 1, y = 70, label = "n = 60", size = 3)

barplot_allneuro
barplot_allneuro <- barplot_allneuro + 
  scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125, 150, 175, 200), labels = c(0, 25, 50, 75, 100, 125, 150, 175, 200))

```

Make barplot of all conditions except stroke
```{r}

barplot_data_notstroke <-
  barplot_data %>% 
  mutate(
    any_figure_contains = case_when(
     spbg_present == "Yes" ~ "SPBG", 
     figure_other == "Yes" & spbg_present == "No" ~ "Other figure (No SPBG)", 
     figure_other == "No" & spbg_present == "No" ~ "No figure")) %>% 
  filter(neuropathology!= "Stroke") %>% 
  mutate(
    neuropathology = recode(neuropathology,
      "TBI" = "Traumatic brain injury"
    )
  )



barplot_data_notstroke$neuropathology[barplot_data_notstroke$neuropathology%in% names(which(table(barplot_data_notstroke$neuropathology) < 3))] = "Other †"

positions_notstroke <- c("Other †", "Traumatic brain injury", "Guillain-Barré",  "Encephalitis", "Huntington’s disease", "Parkinson's disease", "Multiple sclerosis")


barplot_notstroke <- ggplot(barplot_data_notstroke, aes(x = neuropathology, fill = any_figure_contains)) + 
  geom_bar(width = 0.8) +
  scale_fill_manual(values = c("grey93", "#FFE082", "#a4c3d7")) + 
  scale_x_discrete(limits = positions_notstroke) +
  coord_flip() +
  theme(
    panel.background = element_blank(), 
    panel.grid.major.x = element_line(color = "lightgrey",
                                          size = 0.25,
                                          linetype = 5), 
     panel.grid.minor.x = element_line(color = "lightgrey",
                                          size = 0.5,
                                          linetype = 3),
    axis.text.x = element_text(size= 9, face="bold", colour = "black"), 
    axis.text.y = element_text(size= 9, colour = "black"), 
    axis.title.y=element_blank(), 
    legend.title=element_blank(), 
    legend.position = "none",
    axis.ticks = element_blank(), 
    panel.border = element_rect(colour = "lightgrey", fill=NA, size=0.5), 
    axis.title.x = element_text(size = 10), 
    plot.margin = unit(c(1.0,0,0,0), "cm")) + 
  ylab("Number of Papers") +
  geom_text(stat= 'count' , aes(label=..count..), 
            position = position_stack(vjust = .5), 
            color = 'grey16', 
            size = 2.5) +
  ylim(0, 27) +
  annotate("text", x = 1, y = 11.5, label = "n = 10", size = 3)+
  annotate("text", x = 2, y = 5.5, label = "n = 4", size = 3) +
  annotate("text", x = 3, y = 4.5, label = "n = 3", size = 3) +
  annotate("text", x = 4, y = 4.5, label = "n = 3", size = 3) +
  annotate("text", x = 5, y = 4.5, label = "n = 3", size = 3) +
  annotate("text", x = 6, y = 13.5, label = "n = 12", size = 3) +
  annotate("text", x = 7, y = 26.5, label = "n = 25", size = 3)


barplot_notstroke

barplot_notstroke <- 
  barplot_notstroke +
  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 25), labels = c(0, 5, 10, 15, 20, 25))

 barplot_notstroke
    
```
("lightcyan2", "#FFE082", "darkorange2")) 
 device= 'tiff', 
       dpi=700)
       ggsave("figure3_draft4.png")
Combine both bargraphs into one figure
```{r}
combined_barplot <- ggarrange(
  barplot_allneuro,
  barplot_notstroke, 
  labels = c("A", "B"),
  ncol = 1, nrow = 2, 
  heights  = c(1.23,2), 
  align = "v")

combined_barplot

ggsave("figure3_draft5.png")
```


```{r}

gA <- ggplotGrob(barplot_allneuro)
gB <- ggplotGrob(barplot_notstroke)
gB$layout$clip[gB$layout$name == "panel"] <- "off"
maxWidth = grid::unit.pmax(gA$widths[2:5], gB$widths[2:5])
gA$widths[2:5] <- as.list(maxWidth)
gB$widths[2:5] <- as.list(maxWidth)

gB$widths <- gA$widths

grid.arrange(gA, gB, ncol=1)

```

