---
title: "Untitled"
author: "Meghan Forrest"
date: "2023-04-11"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(stringr)
library(ggplot2)
library(readxl)
library(table1)
library(qwraps2)
library(ggthemes)
library(ggpubr)
library(gridExtra)
library(datasets)
library(forcats)
library(patchwork)
library(grid)
library(table1)

Abstraction_master <- read.csv(file = url("https://raw.githubusercontent.com/meghanrforrest/SPBG-meta-research-study/main/Abstraction_master.csv"))
```

# Import dataset and restructure.

```{r}

# filter for only included studies

dat_incl <- Abstraction_master |> 
  filter(eligiblity == "Yes") |> 
  
  
  # create variable for spbg count per study  
  
  mutate(
    spbg_count = case_when(
      spbg_present == "No" ~ 0, 
      spbg_present == "Yes" & additional_spbg_1 == "No" ~ 1, 
      spbg_present == "Yes" & additional_spbg_1 == "Yes" & additional_spbg_2 == "No" ~ 2, 
      spbg_present == "Yes" & additional_spbg_1 == "Yes" & additional_spbg_2 == "Yes"  & additional_spbg_3 == "No" ~ 3, 
      spbg_present == "Yes" & additional_spbg_1 == "Yes" & additional_spbg_2 == "Yes" & additional_spbg_3 == "Yes" & additional_spbg_4 == "No" ~ 4, 
      spbg_present == "Yes" & additional_spbg_1 == "Yes" & additional_spbg_2 == "Yes" & additional_spbg_3 == "Yes" & additional_spbg_4 == "Yes" ~ 5)) |>
  
  # create variable for records which contain an ADJUSTED spbg
  mutate(
    contains_adj_spbg = case_when(
      spbg_present == "Yes" & c(adjusted_1 == "Yes" | adjusted_2 == "Yes" | adjusted_3 == "Yes" | adjusted_4 == "Yes" | adjusted_5 == "Yes") ~ "Yes", 
      spbg_present == "No" ~ NA, 
      spbg_count == 1 & adjusted_1 == "No" ~ "No",
      spbg_count == 2 & c(adjusted_1 == "No" & adjusted_2 == "No") ~ "No", 
      spbg_count == 3 & c(adjusted_1 == "No" & adjusted_2 == "No" & adjusted_3 == "No") ~ "No",
      spbg_count == 4 & c(adjusted_1 == "No" & adjusted_2 == "No" & adjusted_3 == "No" & adjusted_4 == "No") ~ "No",
      spbg_count == 5 & c(adjusted_1 == "No" & adjusted_2 == "No" & adjusted_3 == "No" & adjusted_4 == "No" & adjusted_5 == "No") ~ "No")
    ) |> 
  

  
  # separate out neurological conditions by ;
  separate(neurological_condition, into = c("condition1", "condition2"),
             sep = ";",   
             fill = "right",   
             remove = TRUE) |> 
  
  # clean up names of neuropathologies
  mutate(condition1 = recode(condition1,
      "Stroke (all)" = "Stroke (all)", 
      "encephalitis" = "Encephalitis", 
      "GBS" = "Guillain-Barré", 
      "Huntington disease" = "Huntington’s disease", 
      "Huntington Disease" = "Huntington’s disease")
      ) |> 
  
  mutate(condition2 = if_else( 
         condition2 == "", NA_character_, condition2)
         )|> 
  
  # separate out functional outcomes by ;
  separate(functional_outcome, into = c("outcome1", "outcome2", "outcome3"),
             sep = ";",   
             fill = "right",   
             remove = FALSE) |> 
  
  # standardize formatting of functional outcomes
  mutate(
    outcome1 = recode(outcome1, 
                         '(KOSCHI) King’s Outcome Scale for Childhood Head Injury score' = 'KOSCHI', 
                         'Bayley Scales of Infant Development III' = 'Bayley III', 
                         'Pediatric Stroke Outcome Measure (PSOM) ' = 'PSOM', 
                         'Scales  for  Outcomes  in  PD  (SCOPA)' = "SCOPA", 
                         "Unified  Huntington  Disease  Rating  Scale" = "UHDRS", 
                         "UPDRS" = "UPDRS (composite)", 
                         "Total motor scale (TMS)" = "TMS",
                         "medical research council ( MRC) sum score" = "MRC sum score",
                         "GBS disability score" = "GBS disability scale", 
                         "GBS  disability  scale" = "GBS disability scale",
                         "Pediatric Stroke Outcome Measure (PSOM)" = "PSOM", 
                         "GOS" = "GOS (unextended)")
  )  |>     
  mutate(
    outcome2 = recode(outcome2, 
                      " Gross Motor Function Classification System (GMFCS)" = "GMFCS", 
                      "UPDRS-III" = "UPDRS-III", 
                      "MSSS" = "MSSS", 
                      "mRS" = "mRS", 
                      "NIHSS" = "NIHSS", 
                      " PSOM" = "PSOM", 
                      " total functional capacity (TFC)" = "TFC", 
                      " UPDRS-III" = "UPDRS-III", 
                      "Barthel Index" = "Barthel Index", 
                      "ICARS" = "ICARS", 
                      "mNIHSS" = "mNIHSS", 
                      " NIHSS" = " NIHSS", 
                      " Katz index" = "Katz index", 
                      " Rasch- Built  Overall Built Disability Score" = "Rasch-Built Overall Built Disability Score", 
                      " medical research council ( MRC) sum score" = "Medical Research Council (MRC) sum score", 
                      " GDS" = "GDS", 
                      "Multiple Sclerosis Severity Score (MSSS)" = "MSSS", 
                      " PCPC" = " PCPC", 
                      "Stroke Levity Scale" = "Stroke Levity Scale", 
                      " SCIM 3" = "SCIM 3"),
    outcome2 = if_else(
      outcome2 == "", NA_character_, outcome2)
    ) |> 
  
  mutate(
    outcome3 = recode(outcome3,
                      "INAS" = "INAS", 
                      "Stroke Impact Scale 16" = "Stroke Impact Scale 16",
                      "SIS-16" = "Stroke Impact Scale 16"), 
     outcome3 = if_else(
      outcome3 == "", NA_character_, outcome3)
    ) |> 
  
  # standardize formatting of roc/auc variable
  mutate(
    roc_present = str_replace_all(
      roc_present, ";", "")
  ) |> 
  
  mutate(
    roc_present = recode(
      roc_present,
      "AUC" = "auc", 
      "AUCROC" = "auc_roc", 
      "ROCAUC" = "auc_roc", 
      "Neither" = "neither")
    ) |> 
  
  separate(
    effect_estimate, 
    into = c("effectmeasure1", 
             "effectmeasure2", 
             "effectmeasure3"),
             sep = ";",   
             fill = "right",   
             remove = FALSE) |> 
  # standardize formatting of effect estimate variable
  mutate(
    effectmeasure1 = recode(
      effectmeasure1, 
      "partial correlation r" = "Partial correlation r", 
      "pearson correlation r" = "Pearson correlation r"), 
    
    effectmeasure2 = str_remove(effectmeasure2, "^ "), 
    effectmeasure2 = recode(
      effectmeasure2, 
      "risk ratio" = "Risk ratio", 
      "odds ratio" = "Odds ratio"), 
    effectmeasure2 = if_else(
      effectmeasure2 == "", NA_character_, effectmeasure2),
    
    effectmeasure3 = str_remove(effectmeasure3, "^ "), 
    effectmeasure3 = if_else(
      effectmeasure3 == "", NA_character_, effectmeasure3),
    effectmeasure3 = recode(
      effectmeasure3, 
      "mean difference" = "Mean difference") 
  ) |> 
  
  # create a binary variable for whether a row only contains a beta coefficient
    mutate(
    beta_only_bin = case_when(
      effect_estimate == "Beta coefficient(s) only" ~ "Yes", 
      effect_estimate == "Beta coefficient(s) only;" ~ "Yes",
      TRUE ~ "No")
  ) |> 
  
  # create a binary variable for whether an observation is in the primary sample or only overall sample
  mutate(
    refined_bin = case_when(
      roc_present == "neither" & covariate_intent == "Yes" & beta_only_bin == "No" ~ "refined",
      TRUE ~ "overall_only")
  ) |> 
  
  mutate(
    spbg_type_3cat = case_when(
      spbg_present == "No" ~ "No Grotta bars",
      contains_adj_spbg == "No" ~ "Unadjusted Grotta bars only", 
      contains_adj_spbg == "Yes" ~ "Contains adjusted Grotta bars"),
    spbg_type_3cat = fct_relevel(
      spbg_type_3cat, 
      "No Grotta bars", 
      "Unadjusted Grotta bars only", 
      "Contains adjusted Grotta bars")
  ) |> 
  

  # remove unneeded variables
  select(c("doi",
           "condition1",
           "condition2", 
           "outcome1", 
           "outcome2", 
           "outcome3",
           "spbg_present", 
           "contains_adj_spbg", 
           "spbg_type_3cat", 
           "figure_other",
           "roc_present", 
           "covariate_intent", 
           "effectmeasure1", 
           "effectmeasure2", 
           "effectmeasure3", 
           "beta_only_bin", 
           "refined_bin")
         )

```

# Generate tables

### Supplementary Table 1

Table of neurological conditions in study populations

```{r}
# prepare dataset for neurological conditions

dat_supptab1_conditions <- dat_incl |> 
  pivot_longer(
    cols = starts_with("condition"), 
    names_to = "which_condition", 
    values_to = "condition"
  ) |> 
  mutate(
    refined_bin = recode(
      refined_bin,
      "overall_only" = "Overall", 
      "refined" = "Refined"), 
    
    condition = recode(
      condition, 
      "acute myocardial infarction (AMI)" = "Acute myocardial infarction *", 
      "spinocerebellar ataxia type 2 (SCA2)" = "Spinocerebellar ataxia type 2", 
      "Inclusion body myositis (IBM)" = "Inclusion body myositis", 
      "TBI" = "Traumatic brain injury †",
      "Spinal Cord Injury" = "Spinal cord injury", 
      "Hemorrhagic stroke" = "Stroke (hemorrhagic) †", 
      "Ischemic stroke" = "Stroke (ischemic) *", 
      "Stroke (all)" = "Stroke (hemorrhagic and ischemic)")
    ) |> 
  
  filter(!is.na(condition)) |> 
  
  dplyr::rename(`Neurological condition` = `condition`) 

footnote1 <- "* One occurrence of acute myocardial infarction and ischemic stroke in a combined population"
footnote2 <- "† One occurrence of stroke (hemorrhagic) and traumatic brain injury in a combined population" 

my.render.cat <- function(x) {
  c("", sapply(stats.default(x), function(y) with(y,
                                                  sprintf("%d (%0.0f%%)", FREQ, PCT))))
}

supp_tab1 <- table1(~ `Neurological condition` | spbg_type_3cat, data = dat_supptab1_conditions, 
                    overall=c(left="Total"), 
                    footnote = c(footnote1, footnote2), 
                    render.categorical=my.render.cat)

supp_tab1

```

### Supplementary Table 2

Table of functional outcomes identified in sample

```{r}
dat_supptab2_outcomes <- dat_incl |> 
  pivot_longer(
    cols = starts_with("outcome"), 
    names_to = "which_outcome", 
    values_to = "outcome") |> 
  mutate(
    outcome = recode(
      outcome,
      "mRS" = "Modified Rankin Scale", 
      "GMFCS" = "Gross Motor Function Classification System", 
      "UPDRS (composite)" = "Unified Parkinson's Disease Rating Scale (composite)", 
      "UPDRS-III" = "Unified Parkinson's Disease Rating Scale-III", 
      "NIHSS" = "National Institutes of Health Stroke Scale", 
      "EDSS" = "Expanded Disability Status Scale", 
      "MSSS" = "Multiple Sclerosis Severity Score", 
      "PedNIHSS" = "National Institutes of Health Stroke Scale (pediatric)", 
      "PSOM" = "Pediatric Stroke Outcome Measure", 
      "TMS" = "The Unified Huntington's Disease Rating Scale ‐ Total Motor Score",
      "TFC" = "The Unified Huntington's Disease Rating Scale ‐ Total Functional Capacity",
      "SARA" = "Scale for Assessment and Rating of Ataxia",
      "ICARS" = "International Cooperative Ataxia Rating Scale",
      "INAS" = "Inventory of Non-Ataxia Signs",
      "RS" = "Rankin Scale",
      "mNIHSS" = "Modified National Institutes of Health Stroke Scale",
      "GOS (unextended)" = "Glasgow Outcome Scale",
      " NIHSS" = "National Institutes of Health Stroke Scale",
      "Katz index" = "Katz Index",
      "GOSE" = "Glasgow Outcome Scale - Extended",
      "GBS disability scale" = "Guillain-Barré Syndrome Disability Scale",
      "Medical Research Council (MRC) sum score" = "Medical Research Council Sum Score",
      "I-RODS" = "Inflammatory Rasch-Built Overall Disability Scale",
      "GDS" = "Guillain-Barré Syndrome Disability Scale",
      "WIDEA-FS" = "The Warner Initial Developmental Evaluation of Adaptive and Functional Skills",
      "KOSCHI" = "King's Outcome Scale for Childhood Head Injury",
      "UHDRS" = "Unified Huntington's Disease Rating Scale",
      "SCOPA" = "Scales for Outcomes in Parkinson's Disease",
      "GOS-E-Peds" = "Glasgow Outcome Scale - Extended (pediatric)",
      " PCPC" = "Pediatric Cerebral Performance Category Scale",
      "IBMFRS" = "Inclusion Body Myositis-Functional Rating Scale",
      "UW-mRS" = "Utility-Weighted Modified Rankin Scale",
      "ISNCSCI" = "International Standards for Neurological Classification of Spinal Cord Injury",
      "SCIM 3" = "Spinal Cord Independence Measure, Version III",
      "Ped-MSSS" = "Multiple Sclerosis Severity Score (pediatric)")
  ) |> 
  filter(!is.na(outcome)) |> 
  
  dplyr::rename(`Functional outcome` = `outcome`) 
  


my.render.cat <- function(x) {
  c("", sapply(stats.default(x), function(y) with(y,
                                                  sprintf("%d (%0.0f%%)", FREQ, PCT))))
}

supp_tab2 <- table1(~ `Functional outcome` | spbg_type_3cat, data = dat_supptab2_outcomes, 
                    overall=c(left="Total"), 
                    render.categorical=my.render.cat)

supp_tab2
```

# Generate barplots

#### Prepare a dataset for barplots

```{r}
dat <- Abstraction_master |> 
  mutate(
    any_figure_contains = case_when(
     spbg_present == "Yes" & contains_adj_spbg == "No" ~ "Unadjusted SPBG", 
     contains_adj_spbg == "Yes" ~ "Adjusted SPBG", 
     figure_other == "Yes" & spbg_present == "No" ~ "Other figure (No SPBG)", 
     figure_other == "No" & spbg_present == "No" ~ "No figure"),
    
  # brush up the names of the neurological conditions
    condition1 = recode(
      condition1,
      "Ischemic stroke" = "Stroke", 
      "Hemorrhagic stroke" = "Stroke", 
      "Stroke (all)" = "Stroke"
    )) |> 
  
  # make the records with two populations into their own observation
  pivot_longer(
    names_to = "which_condition", 
    values_to = "neuropathology", 
    cols = condition1:condition2) %>% 
  filter(neuropathology != "") |> 
  select(!"which_condition") |> 
  
  # clean up names of neuropathologies
  mutate(neuropathology = recode(neuropathology,
      "Stroke (all)" = "Stroke (all)", 
      "encephalitis" = "Encephalitis", 
      "GBS" = "Guillain-Barré", 
      "Huntington disease" = "Huntington’s disease", 
      "Huntington Disease" = "Huntington’s disease")
      ) |>
  
  # relevel the any_figure_contains variable
  mutate(
    any_figure_contains = fct_relevel(any_figure_contains,
                                      "No figure", 
                                      "Other figure (No SPBG)", 
                                      "Unadjusted SPBG",
                                      "Adjusted SPBG"), 
    any_figure_contains = recode(any_figure_contains, 
                                 "SPBG" = "Stacked proportional bar graph"))
```

mutate( any_figure_contains = recode( any_figure_contains, "SPBG" =
"Stacked proportional bar graph"))

```{r}

#set up a dataframe for the "Stroke v. Other" barplot
dat_barplot_stroke <- dat |> 
  mutate(
    neuropathology = fct_other(neuropathology, keep = "Stroke", 
                                    other_level = "Other *"), 
    neuropathology = fct_relevel(neuropathology, "Other *", "Stroke"))



barplot_stroke <- ggplot(dat_barplot_stroke, 
                         aes(x= neuropathology, 
                             fill = any_figure_contains)) + 
  geom_bar(width = 0.5, 
           color = "white") + 
  
  coord_flip() +
  
  scale_fill_manual(values = c("#533E57", 
  "#8b738e", 
  "#d2b3d6",
  "#d9f4e6"), 
                    limits = c("Adjusted SPBG", 
                               "Unadjusted SPBG", 
                               "Other figure (No SPBG)",
                               "No figure")) +
  
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(), 
    panel.grid.minor.y = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.text.y = element_blank(),
    
    legend.title=element_blank(),
    legend.key.size = unit(1, 'cm'),
        legend.key.height = unit(.3, 'cm'),
        legend.key.width = unit(.35, 'cm'),
        legend.text = element_text(size = 9), 
    legend.position = c(0.5, 1.0),
    legend.direction="horizontal",
    
    axis.ticks = element_blank(), 
    
    axis.title.y=element_blank(), 
    axis.title.x=element_blank(), 
    

    plot.margin = unit(c(0.8,3,0,0), "cm")) +
  geom_text(
    stat= 'count' , aes(label=..count..), 
            position = position_stack(vjust = .5), 
            size = 3, 
            color = 'black') +
  annotate("text", x = 2, y = 215, label = "Total = 192", size = 3, fontface = "bold"
           ) +
  annotate("text", x = 1, y = 83, label = "Total = 60", size = 3, fontface = "bold") +
  annotate("text", x = 2.35, y = 106, label = "Studies in the sample with a stroke population", size = 4, fontface = "bold") +
  annotate("text", x = 1.35, y = 116, label = "Studies in the sample with a non-stroke population", size = 4, fontface = "bold") +
  scale_y_continuous(limits = c(0, 250))

barplot_stroke



```

barplot_stroke barplot_allneuro \<- barplot_allneuro +
scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125, 150, 175, 200),
labels = c(0, 25, 50, 75, 100, 125, 150, 175, 200))

Make barplot of all conditions except stroke

```{r}

barplot_data_notstroke <-
  dat %>% 
  mutate(
    any_figure_contains = case_when(
     spbg_present == "Yes" ~ "SPBG", 
     figure_other == "Yes" & spbg_present == "No" ~ "Other figure (No SPBG)", 
     figure_other == "No" & spbg_present == "No" ~ "No figure")) |> 
  filter(neuropathology!= "Stroke") |> 
  mutate(
    neuropathology = recode(neuropathology,
      "TBI" = "Traumatic brain injury"
    )
  )



barplot_data_notstroke$neuropathology[barplot_data_notstroke$neuropathology%in% names(which(table(barplot_data_notstroke$neuropathology) < 3))] = "Other †"

positions_notstroke <- c("Other †", 
                         "Traumatic brain injury", 
                         "Guillain-Barré",  
                         "Encephalitis", 
                         "Huntington’s disease", 
                         "Parkinson's disease", 
                         "Multiple sclerosis")


barplot_notstroke <- ggplot(barplot_data_notstroke, aes(x = neuropathology, fill = any_figure_contains)) + 
  geom_bar(width = 0.9, color = "white") +
  scale_fill_manual(values = c("#d9f4e6", 
                               "#d2b3d6", 
                                "#8b738e")) + 


  scale_x_discrete(limits = positions_notstroke) +
  coord_flip() +
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(),  
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.background = element_blank(), 
    axis.text.y = element_text(size= 9, colour = "black"), 
    axis.title.y=element_blank(), 
    legend.title=element_blank(), 
    legend.position = "none",
    axis.ticks = element_blank(), 
    axis.title.x = element_text(size = 10), 
    plot.margin = unit(c(1.0,0,0,0), "cm")) +
  
  ylab("Number of Papers") +
  geom_text(stat= 'count' , aes(label=..count..), 
            position = position_stack(vjust = .5), 
            color = 'grey16', 
            size = 2.5) +

  annotate("text", x = 1, y = 13.5, label = "Total = 10", size = 3)+
  annotate("text", x = 2, y = 8.5, label = "Total = 4", size = 3) +
  annotate("text", x = 3, y = 7.5, label = "Total = 3", size = 3) +
  annotate("text", x = 4, y = 7.5, label = "Total = 3", size = 3) +
  annotate("text", x = 5, y = 7.5, label = "Total = 3", size = 3) +
  annotate("text", x = 6, y = 16.5, label = "Total = 12", size = 3) +
  annotate("text", x = 7, y = 29.5, label = "Total = 25", size = 3) +
  
  scale_y_continuous(limits = c(0, 50))


barplot_notstroke

```

("lightcyan2", "#FFE082", "darkorange2")) device= 'tiff', dpi=700)
ggsave("figure3_draft4.png") Combine both bargraphs into one figure

```{r}
combined_barplot <- ggarrange(
  barplot_stroke,
  barplot_notstroke, 
  labels = c("A", "B"),
  ncol = 1, nrow = 2, 
  heights  = c(1.23,2), 
  align = "v")

combined_barplot

#ggsave("figure3_draft5.png")
```

```{r}

gA <- ggplotGrob(barplot_stroke)
gB <- ggplotGrob(barplot_notstroke)
gB$layout$clip[gB$layout$name == "panel"] <- "off"
maxWidth = grid::unit.pmax(gA$widths[2:5], gB$widths[2:5])
gA$widths[2:5] <- as.list(maxWidth)
gB$widths[2:5] <- as.list(maxWidth)

gB$widths <- gA$widths

combined_plot1 <- grid.arrange(gA, gB, ncol=1) 
grid.lines(x = unit(c(0.34, 0.8), "npc"), y = unit(c(0.51, 0.42), "npc"), gp = gpar(col = "darkgrey"))
grid.text("Breakdown of non-stroke populations in sample", x = unit(0.05, "npc"), y = unit(0.44, "npc"), just = "left", gp = gpar(fontsize = 10, fontface = "bold"))





```
